---
title: "Statistical Inference Course Project Part 2"
author: "Sarthak"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r, results='hold'}
# loading the required libraries
library(ggplot2)
library(dplyr)
```

# Basic Inferential Data Analysis

## 1. Overview

In this analysis, the ToothGrowth data will be analyzed. This shows the effect of vitamin C on teeth-growth in guinea pigs. Two vitamin C supplements are used, with varying dose levels.

## 2. Loading the Dataset and basic Exploratory Data Analysis

The ToothGrowth data set has to be loaded from the `datasets` package in R.

```{r}
# loading the datasets package
library(datasets)

# loading the ToothGrowth data set
data("ToothGrowth")
```

Next, a plot will be created showing the trends in tooth-growth with respect to each supplement

```{r}
# creating a point plot of the variation of teeth growth with respect to each dose level 
# of a supplement, and fitting a linear line over those points
qplot(x = dose, y = len, data = ToothGrowth, color = supp, geom = "point",
      main = "Tooth Growth with each Dose Level of a Supplement",
      xlab = "Dose Level", ylab = "Growth in Tooth Length") +
geom_smooth(method = "lm") + theme_bw() + labs(colour = "Supplement")
```

From the plot, it is clear that the tooth length increases with increasing levels of dose of both supplements, and the _OJ_ supplement has higher tooth length growths than the _VC_ supplement. The growth rate for the two supplements merge at dose levels of 2 mg/day, whereas there is some difference in growth levels for the lower dose levels of the two supplements. Also, the growth rate with respect to each dose level of _VC_ is higher than that of _OC_.

## 3. Basic Data Summary

A summary of the whole dataset has been shown below

```{r}
# generating a summary of all variables in the dataset
summary(ToothGrowth)
```

```{r}
# mean tooth growth length
mean(ToothGrowth$len)
```

```{r}
# standard deviation of the length of tooth growth
sd(ToothGrowth$len)
```
